---
title: "London Marathon"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggflags)
```

I started by loading in and exploring the `winners` data set.
I decided that I wanted to highlight the nationalities in the chart, while also highlighting the fastest times in both categories.
```{r}
winners <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-25/winners.csv')
```

 `!grepl` called within the filter function will filter out any rows that contain a specific string within the category column in this example.
 I also changed all countries to their ISO 3166-2 code, note lower case works best for use with `ggflags`
```{r}
WinnersTidy <- winners %>%
  filter(!grepl('Wheelchair', Category)) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'United States' , 'us')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Norway' , 'no')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'United Kingdom' , 'gb')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Japan' , 'jp')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Denmark' , 'dk')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Kenya' , 'ke')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Soviet Union' , 'ru')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Portugal' , 'pt')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Mexico' , 'mx')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Spain' , 'es')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Morocco' , 'ma')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Ethiopia' , 'et')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Poland' , 'pl')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Germany' , 'de')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'Ireland' , 'ie')) %>%
   mutate(Nationality = replace(Nationality, Nationality == 'China' , 'cn'))

```
 
I decided to split both categories into a facet grid, as men and women compete separately I did not want them on the same axis. I set the scales to free which allows both y axis to adjust to fit each set of data better.
After a lot of trial and error I managed to get `geom_flag()` working, geom flag hits an error when looking for a Soviet Union flag, there is a work around available for this, but in this case I just used the Russia flag instead.

```{r  fig.retina = 2}
  ggplot(WinnersTidy, aes(x = Year, y = Time)) +
  facet_grid(rows = vars(Category), scales = "free") +
  geom_flag(aes(country = Nationality), size = 5) 
```

I have now decided that I want to label the course records with their name and time, so I now need to extract these from the data set and create a new data set RecordsTidy.
```{r}
RecordsTidy <- WinnersTidy %>% 
   group_by(Category) %>% 
   slice_min(order_by = Time)

RecordsTidy
```
In order to combine both data sets into the chart I needed to reorder some of the ggplot call to load the relevant data set into each respective geom. This allows me to have only the RecordsTidy subset data labelled rather than each flag. Also I could not manually add the text to the chart, im not sure if it is because of the facet or becuase of the time data being difficult to work with.

```{r  fig.retina = 2}
  ggplot() +
  geom_flag(data = WinnersTidy,aes(x = Year, y = Time, country = Nationality), size = 5) +
  geom_text(data = RecordsTidy, aes(x = Year, y = Time, label = Athlete),
            hjust = +1.1, size = 3) +
  geom_text(data = RecordsTidy, aes(x = Year, y = Time, label = Time),
            hjust = -0.2, size = 3) +
  facet_grid(rows = vars(Category), scales = "free")+
  theme_minimal() +
  labs(title = "Nationality of London Marathon Winners by Year ",
       subtitle = "Kipchoge and Radcliffe hold the course record \nfor the Men and Women respectively",
       caption = "Twitter: @Louis_Caruana")

```


